Secure Composition of Robust and Optimising Compilers

### 2023-07-05 popl/main.tex
edits to entry to 2

\section{Background\pages{1}}\label{sec:background}
To reason about security, this section makes the concepts of properties and the satisfaction thereof precise (\Cref{subsec:bg:tprop}).
Furthermore, the section introduces robustness in the context of both satisfaction and secure compilers (\Cref{subsec:bg:rtp}) in similar fashion to existing work~\cite{abate2019jour}.

\subsection{Properties, Satisfaction, and Robustness}\label{subsec:bg:tprop}

The intuition behind security is that nothing bad happens.
Properties ($\pi$) make this intuition formally precise, however, they rely on a trace model.
{\em Traces} ($\trace$) are (possibly infinite) sequences of {\em events} ($\event$), which are an abstract representation of observable actions.
The empty event ($\emptyevent$) is an uboservable or unimportant action.
Throughout the paper, empty sequences are denoted with $\hole{\cdot}$ and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ with $\trace[_{1}]\cdot\trace[_{2}]$.
For prepending events to traces the paper uses the same notation, i.e., $\event\cdot\trace$.

A property $\pi$ is a set of admissable traces and $\pi$ is a security property if a finite prefix of a trace $\trace$ is enough to decide wether $\trace\in\pi$ or $\trace\notin\pi$.
For example, given the trace $\ev{Send\ 1}\cdot\ev{Recv\ 1}\cdot\ev{Send\ 2}\cdot\dots$ and the property $\pi=\left\{\trace \mid \trace=\hole{\cdot} \text{ or } \trace=\ev{Send\ n}\cdot\ev{Recv\ 2n}\cdot\trace[']\wedge \trace[']\in\pi\right\}$, the prefix $\ev{Send\ 1}\cdot\ev{Recv\ 1}$ immediately violates $\pi$.
The set of security properties is one of many classes of properties ($\class$).
Note that existing literature refers to classes of properties as hyperproperties~\cite{clarkson2008hyper}.
The framework developed in this paper relies on {\em liftings} ($\lift{\pi}$), which are simply the powerset of a property $\pi$ and thus also considered as a class.
This framework is concerned about the secure compilation of programs that admit properties.

A program ($\runtimetermvar$) is a tuple consisting of some dynamic state ($\statevar$) and a syntactic description ($\progvar$).
Programs without free variables are considered {\em whole}.
Contrary, they are {\em partial} if they contain variables without an accompanying definition.
Whole programs can be run with the multi-step relation $\stepto{\runtimetermvar}{\runtimetermvar[']}{\trace}$ and yield a trace $\trace$.
Moreover, the paper assumes for such executions that $\runtimetermvar[']$ is {\em final}, decided by the judgement $\isval{\runtimetermvar}$.
Partial programs cannot be executed, since execution would get stuck when reaching an undefined variable.
The paper also uses other step-relations, such as a single-step ($\stepto[]{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\event}$) and an $n$-step relation ($\stepton{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\trace}$) that performs $n$ single-steps.
Note that the multi-step relation presented earlier is defined as-standard using the single-step, where the reflexive case makes use of $\isval{\runtimetermvar}$.
All multi-step relations filter $\emptyevent$ events.

The {\em behavior} of a whole program $\runtimetermvar$ ($\behav{\runtimetermvar}$) is a set containing all traces the program may emit during execution.
%This paper only considers deterministic languages, where the behavior $\behav{\runtimetermvar}$ is a singleton set for any whole $\runtimetermvar$.
This set is used to define the satisfaction of a property as follows:

\begin{definition}[Satisfaction]\label{def:propsat}
  A whole program $\runtimetermvar$ satisfies $\pi$ ($\sat{\runtimetermvar}{\pi}$) iff $\behav{\runtimetermvar}\in\pi$.
\end{definition}

@@

\begin{definition}[Robust Satisfaction]\label{def:proprsat}
  A program $\runtimetermvar$ robustly satisfies $\pi$ ($\rsat{\runtimetermvar}{\pi}$) iff for any partial program $\contextvar$ it holds that $\sat{\link{\contextvar}{\runtimetermvar}}{\pi}$.
\end{definition}

\subsection{Secure Compilers}\label{subsec:bg:rtp}

>>>>

\section{Background: Security Properties and Secure Compilers\pages{1}}\label{sec:background}
To introduce the security argument of this paper, this section first presents the concepts of (security) properties, of their satisfaction, and of their robust satisfaction (i.e., satisfaction in the presence of an active attacker; \Cref{subsec:bg:tprop}).
Then, borrowing from existing work~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, the section introduces secure compilers as compilers that preserve robust property satisfaction (\Cref{subsec:bg:rtp}).

\subsection{Properties, Satisfaction, and Robust Satisfaction}\label{subsec:bg:tprop}

% The intuition behind security is that nothing bad happens.
% \MPin{
% 	no: that is safety
% }
This paper employs the security model where programs are written in a language whose semantics emits events $\event$.
Events include security-relevant actions (e.g., reading ad writing to memory) and the unobservable event $\emptyevent$.
As programs executes, their emitted events are concatenated in traces ($\trace$), which serve as the description of the behaviour of a program.%
\footnote{
Throughout the paper, we denote empty sequences with $\hole{\cdot}$ and concatenate sequences $\trace[_{1}],\trace[_{2}]$ as $\trace[_{1}]\cdot\trace[_{2}]$.
Prepending elements to sequences uses the same notation: $\event\cdot\trace$.
}

Properties ($\pi$) are sets of traces of admissible program behaviours, ascribing what said property considers valid.\MP{ cit }

% make this intuition formally precise, however, they rely on a trace model.
% {\em Traces} ($\trace$) are (possibly infinite) sequences of {\em events} ($\event$).
% Events are an abstract representation of observable actions, in this paper
% The empty event ($\emptyevent$) is an unobservable (or unimportant) action.
% A property $\pi$ is a set of admissable traces and 
Properties and partitioned in the subsets of safety and liveness\MP{cit}, and given their central role for security, this paper focusses on safety properties.
Given a safety property, a finite prefix of a longer trace $\trace$ is enough to decide whether a trace is admitted or not by the property. % $\trace\in\pi$ or $\trace\notin\pi$.
For example, given the trace $\ev{Send\ 1}\cdot\ev{Recv\ 1}\cdot\ev{Send\ 2}\cdot\dots$ and the property $\pi=\left\{\trace \mid \trace=\hole{\cdot} \text{ or } \trace=\ev{Send\ n}\cdot\ev{Recv\ 2n}\cdot\trace[']\wedge \trace[']\in\pi\right\}$, the prefix $\ev{Send\ 1}\cdot\ev{Recv\ 1}$ immediately violates $\pi$.
\MP{mention send and recv before already. replace with read/write to contextualise to memory?}

% The set of security properties is one of many classes of properties ($\class$).
% Note that existing literature refers to classes of properties as hyperproperties~\cite{clarkson2008hyper}.
% The framework developed in this paper relies on {\em liftings} ($\lift{\pi}$), which are simply the powerset of a property $\pi$ and thus also considered as a class.
% This framework is concerned about the secure compilation of programs that admit properties.
\MPin{
	this par does not flow.
	what is the point?
}

% A program ($\runtimetermvar$) is a tuple consisting of some dynamic state ($\statevar$) and a syntactic description ($\progvar$).
% Programs without free variables are considered {\em whole}.
% Contrary, they are {\em partial} if they contain variables without an accompanying definition.
% Whole programs can be run with the multi-step relation $\stepto{\runtimetermvar}{\runtimetermvar[']}{\trace}$ and yield a trace $\trace$.
% Moreover, the paper assumes for such executions that $\runtimetermvar[']$ is {\em final}, decided by the judgement $\isval{\runtimetermvar}$.
% Partial programs cannot be executed, since execution would get stuck when reaching an undefined variable.
% The paper also uses other step-relations, such as a single-step ($\stepto[]{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\event}$) and an $n$-step relation ($\stepton{\runtimetermvar[_{1}]}{\runtimetermvar[_{2}]}{\trace}$) that performs $n$ single-steps.
% Note that the multi-step relation presented earlier is defined as-standard using the single-step, where the reflexive case makes use of $\isval{\runtimetermvar}$.
% All multi-step relations filter $\emptyevent$ events.
\MPin{
	all unnecessary.
	unused notation, not needed for sec2, for sec3 and for sec 4.
	also i don't agree with what a program is.
}

Given a whole program $\runtimetermvar$, i.e., \MP{ fill in}, the {\em behaviour} of a whole program $\runtimetermvar$ ($\behav{\runtimetermvar}$) is the set containing all traces the program can emit during execution according to its language semantics.
%This paper only considers deterministic languages, where the behavior $\behav{\runtimetermvar}$ is a singleton set for any whole $\runtimetermvar$.
This set is used to define property satisfaction (\Cref{def:propsat}):

\begin{definition}[Satisfaction]\label{def:propsat}
  A whole program $\runtimetermvar$ satisfies $\pi$ ($\sat{\runtimetermvar}{\pi}$) iff $\behav{\runtimetermvar}\in\pi$.
\end{definition}
\MPin{
	this definition has 1/2 explained in text, 1/2 in formalism.
}
\MP{
	can't we use p for programs?
}

@@

\begin{definition}[Robust Satisfaction]\label{def:proprsat}
  A program $\runtimetermvar$ robustly satisfies $\pi$ ($\rsat{\runtimetermvar}{\pi}$) iff for any partial program $\contextvar$ it holds that $\sat{\link{\contextvar}{\runtimetermvar}}{\pi}$.
\end{definition}
\subsection{Secure Compilers}\label{subsec:bg:rtp}
